---
title: "Analysis"
authors: "Sofía Navarro & Astrid Sletten Rybner"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# set wd to src folder of the cloned repository
setwd() 

# load packages 
library(pacman)
pacman::p_load("dplyr", "stringr", "tidyverse", "sf", "spatstat", "RColorBrewer", "ggplot2", "cowplot", "spatstat", "pryr")

```

# 1. ANALYZING CLUSTERING OF KINGFISHER OBSERVATIONS

Making bird data ready
```{r}
# read preprocessed birds data
kingfisher <- read_sf("../data/data_preprocessed/kingfisher_with_crs.shp")

# subset bird observations per year
s2016 <- kingfisher %>% filter(kingfisher$Year==2016)
s2020 <- kingfisher %>% filter(kingfisher$Year==2020)

# define local crs with utm zone for Denmark (zone 32)
dk_crs = "+proj=utm +zone=32 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

# give the bird data local projected crs
s2016_transformed <- s2016 %>%
  st_transform(crs=dk_crs)
s2020_transformed <- s2020 %>%
  st_transform(crs=dk_crs)

```


Making the borders of DK into an owin object to use for defining PPP object:
```{r}

# read file borders of denmark to use in ppp
borders <- read_sf("../data/data_country_borders/denmark_administrative_boundaries_province_polygon.shp")

# transform crs to DK utm zone 32
borders_transformed <- st_transform(borders, crs = crs(s2016_transformed))

# simplify the borders object 
object_size(borders_transformed)
borders_simple <- st_simplify(borders_transformed, 
                            preserveTopology = TRUE, 
                            dTolerance = 100)
object_size(borders_simple)

# make simplified borders object into owin object
borders_ow <- as.owin(borders_simple, fatal = TRUE)

# plot borders 
plot(borders_ow)

```


Making birds coordinates of 2016 and 2020 into seperate columns so they can be used to create ppp object:
```{r}
# get just the coordinates of the projected birds coordinates
s2016_sf <- as_tibble(st_geometry(s2016_transformed))
s2020_sf <- as_tibble(st_geometry(s2020_transformed))

# seperate coords into two columns
new16 <-  separate(data = s2016_sf, into= c("X", "Y"),col = geometry, sep = ",")
new20 <- separate(data = s2020_sf, into= c("X", "Y"),col = geometry, sep = ",")

# clean collumns
new16_sep  <- new16  %>% 
  mutate(X = as.numeric(str_extract(new16$X, pattern = "\\d+.\\d+")), 
         Y = as.numeric(str_extract(new16$Y, pattern = "\\d+.\\d+")))

new20_sep  <- new20  %>% 
  mutate(X = as.numeric(str_extract(new20$X, pattern = "\\d+.\\d+")), 
         Y = as.numeric(str_extract(new20$Y, pattern = "\\d+.\\d+")))

# make into data.frames
new16_sep <- as.data.frame(new16_sep)
new20_sep <- as.data.frame(new20_sep)

# make a ppp object of birds observations (use min and max longitude and latitude)
birds_ppp_2016 <- ppp(new16_sep[,1], new16_sep[,2], window = borders_ow)
birds_ppp_2020 <- ppp(new20_sep[,1], new20_sep[,2], window = borders_ow)

# plot ppp objects
plot(birds_ppp_2016)
plot(birds_ppp_2020)

```


Compute envelopes of K-function under random locations
```{r}
# compute envelopes 
K_birds_env16 <- envelope(birds_ppp_2016, fun=Kest, nsim=99, rmax= 30000, correction = "border")
K_birds_env20 <- envelope(birds_ppp_2020, fun=Kest, nsim=99, rmax= 30000, correction = "border")


# plot envelopes (insert the full formula to plot K minus pi * r^2)
jpeg(file="../output/env_plot_2016.jpeg", width=550, height=400)
plot(K_birds_env16, . - pi* r^2 ~ r,  main="Monte carlo test: K(r) for 2016")
dev.off()

jpeg(file="../output/env_plot_2020.jpeg", width=550, height=400)
plot(K_birds_env20, . - pi* r^2 ~ r,  main="Monte carlo test: K(r) for 2020")
dev.off()

```


# 2. CORRELATION TEST: WATER QUALITY AND BIRD COUNT

Computing buffers around all lakes:
```{r}
# read pre-processed lakes data
lakes <- read_sf("../data/data_preprocessed/lakes_preprocessed.shp")

# compute the center points (centroids) of the lakes
centroids <- st_centroid(lakes)

# plot the center points of the lakes
plot(st_geometry(lakes))
plot(centroids, pch = 20, col = "firebrick", add = TRUE)

# make buffers around the center points and give local crs
lakes_buffer <- centroids %>% 
  st_transform(crs = dk_crs) %>% 
  st_buffer(dist = 1000)

# limit the object to just geometry (for plotting)
lakes_buffer_geom <- st_geometry(lakes_buffer)

# compute the number of features in lakes_buffer_geom
length(lakes_buffer_geom) # 684 buffers (lakes)

# plot the lakes buffers
plot(lakes_buffer_geom)

# make into sf objects so centroids can be joined with bird observations
s2016_transformed <- st_sf(s2016_transformed)
s2020_transformed <- st_sf(s2020_transformed)
lakes_buffer_geom <- st_sf(lakes_buffer_geom)

# combine centroids and birds in one object 
lakes_birds2016 <- st_join(lakes_buffer_geom, s2016_transformed)
lakes_birds2020 <- st_join(lakes_buffer_geom, s2020_transformed)

# plot the birds in 2016 and 2020 together with buffers
plot(lakes_buffer_geom)
plot(s2016_transformed, add=TRUE, pch = 20, col = "red")
plot(s2020_transformed, add=TRUE, pch = 20, col = "green")



```

Counting bird observations within buffers:
```{r}
# count the number of bird observations within each of the buffers and add this count to the lakes_buffer data frame
lakes_buffer$birdcount2016 <- lengths(st_intersects(lakes_buffer_geom, s2016_transformed))
lakes_buffer$birdcount2020 <- lengths(st_intersects(lakes_buffer_geom, s2020_transformed))

```

Plotting mean bird count per water category in 2016 and 2020:
```{r}

# plot mean bird count per water category 
plot16 <- ggplot(lakes_buffer, aes(x=wtr_qlt, y=birdcount2016))+
  geom_bar(stat='identity', fill="forest green")+
  ylab("Bird count")+
  xlab("Water quality group")+ggtitle("Bird count 2016")+
  theme(axis.title = element_text(size = 20), 
        plot.title = element_text(size = 30),
        axis.text = element_text(size = 20)) 


plot20 <- ggplot(lakes_buffer, aes(x=wtr_qlt, y=birdcount2020))+
  geom_bar(stat='identity', fill="forest green")+
  ylab("Bird count")+
  xlab("Water quality group")+
  ggtitle("Bird count 2020")+
  theme(axis.title = element_text(size = 20), 
        plot.title = element_text(size = 30),
        axis.text = element_text(size = 20)) 

combined <- cowplot::plot_grid(plot16, plot20)

ggsave("bird_count_water_categories.png", plot = combined, path = "../output/", width = 12, height = 6, dpi = 300)

```

Conducting correlation test between bird count and water quality: 
```{r}
# make numeric 
lakes_buffer$wtr_qlt <- as.numeric(lakes_buffer$wtr_qlt)

# correlation test 2016
cor.test(x = lakes_buffer$wtr_qlt, y = lakes_buffer$birdcount2016, method = "pearson")

# correlation test 2020
cor.test(x = lakes_buffer$wtr_qlt, y = lakes_buffer$birdcount2020, method = "pearson")
```

